---
title: "ANLY 545 - Assignment 2"
author: "Laxman Panthi"
output:
  html_document:
    df_print: paged
---

### Exercise 2.3 
The data set UCBAdmissions is a 3-way table of frequencies classified by Admit, Gender, and Dept.
```{r, data = UCBAdmissions}
data("UCBAdmissions")
ftable(UCBAdmissions)
```

(a)	Find the total number of cases contained in this table.
```{r}
# save typing 
sum(UCBAdmissions)
```

(b)	For each department, find the total number of applicants.
```{r}
margin.table(UCBAdmissions, 3)
```


(c)	For each department, find the overall proportion of applicants who were admitted.
```{r}
tab1 <- margin.table(UCBAdmissions , c(1,3)) 

tab1[1,] / margin.table(UCBAdmissions, 3) 
```



(d)	Construct a tabular display of department (rows) and gender (columns), showing the proportion of applicants in each cell who were admitted relative to the total applicants in that cell.
```{r}
#proportion in each cell admitted
admit <- UCBAdmissions[1,,] 
reject <- UCBAdmissions[2,,]
admit / (admit+reject)
```

### Exercise 2.5 
The data set UKSoccer in vcd gives the distributions of number of goals scored by the 20 teams in the 1995/96 season of the Premier League of the UK Football Association.
This two-way table classifies all 20 × 19 = 380 games by the joint outcome (Home, Away), the number of goals scored by the Home and Away teams. The value 4 in this table actually represents 4 or more goals.
```{r}
library(vcd) 
data(UKSoccer)
```

(a)	Verify that the total number of games represented in this table is 380.
```{r}
sum(UKSoccer)
```

(b)	Find the marginal total of the number of goals scored by each of the home and away teams.
```{r}
#marginal totals
(home <- margin.table(UKSoccer, 1))
(away <- margin.table(UKSoccer, 2))
```

(c)	Express each of the marginal totals as proportions.
```{r}
#proportions
(home_prop <- prop.table(margin.table(UKSoccer,1)))
(away_prop <- prop.table(margin.table(UKSoccer,2)))
```

Optional: (d)	Comment on the distribution of the numbers of home-team and away-team goals. Is there any evidence that home teams score more goals on average?

```{r}
summary(UKSoccer)
```
p-value 0.2846 indicates the null hypothesis cannot be rejected. There is no significant difference between the two factors.


