---
Title: "ANLY545 Assignement #4"
Author: "Harsh Patel"
Date: "`r Sys.Date()`"
output: word_document
always_allow_html: yes

---



```{r }


library(MASS)
library(vcd)
library(vcdExtra)
library(logmult)

```


##Exercise 1: 

```{r }

data(HairEyePlace)

# separate mosaics
mosaic(HairEyePlace[,,1], shade=TRUE, main="Caithness")
mosaic(HairEyePlace[,,2], shade=TRUE, main="Aberdeen")

# condition on Place
mosaic(~Hair + Eye |Place, data=HairEyePlace, shade=TRUE, legend=FALSE)
cotabplot(~Hair+Eye|Place, data=HairEyePlace, shade=TRUE, legend=FALSE)

```

```{r }


data("Accident", package = "vcdExtra")
str(Accident)


```

#1) Use loglm() to fit the model of mutual independence, Freq ~ age + mode+ gender + result to this data.


```{r }

library(MASS)

loglm(formula = Freq ~ age + mode + gender + result, data = Accident)

accident.mod <- loglm(Freq ~ age + mode + gender + result, data = Accident)

```

#2) Use mosaic() to produce an interpretable mosaic plot of the associations among all variables under the model of mutual independence.

```{r }

mosaic(Freq ~ age + mode + gender + result, data = Accident, shade = TRUE, labeling_args = list(clip = c(result = TRUE)))

Accident$mode <- ordered(Accident$mode, levels=levels(Accident$mode)[c(4,2,3,1)])
str(Accident)

levels(Accident$mode)


```

#3)Try different orders of the variables in the mosaic. (Hint: the abbreviate component of the labeling_args argument to mosaic() will be useful to avoid some overlap of the category labels.)
```{r }

mosaic(Freq ~ age + mode + gender + result, data = Accident, shade = TRUE, labeling_args = list(clip = c(result = TRUE)))

mosaic (Freq ~ gender + mode  + result + age , data = Accident, shade = TRUE, labeling_args = list(clip = c(result = TRUE)))

mosaic (Freq ~ result + gender + age + mode, data = Accident, shade = TRUE, labeling_args = list(clip = c(result = TRUE)))



```
#4)Treat result ("Died" vs. "Injured") as the response variable, and fit the model Freq ~ age*mode*gender + result that asserts independence of result from all others jointly.

```{r }
accident.mod2 <- loglm(Freq ~ age * mode * gender + result, data = Accident)

accident.mod2

```

#5)Construct a mosaic display for the residual associations in this model. Which combinations of the predictor factors are more likely to result in death?

```{r }

mosaic(accident.mod2, shade = TRUE, labeling = labeling_residuals, rot_labels = c(20, 20, 30, 90))


```


##Exercise #2.

```{r }

data("caith", package = "MASS")
caith

```
#1)The loglm() and mosaic() functions don’t understand data in this format, so use Caith <- as.matrix(caith) to convert to array form. Examine the result and use names(dimnames(Caith))<-c() to assign appropriate names to the row and column dimensions

```{r }

Caith <- as.matrix(caith)
dimnames(Caith)

names(dimnames(Caith)) <- c("Eye", "Hair")
```

#2)Fit the model of independence to the resulting matrix using loglm().
```{r }

(caith.mod <- loglm(~Hair+Eye, data=Caith, fitted=TRUE))

```

#3)Calculate and display the residuals for this model.
```{r }

 residuals(caith.mod)

```
#4)Create a mosaic display for this data.

```{r }

mosaic(Caith, shade=TRUE)


```




